Understanding Subsidy Allocation in Toronto's Child Care Centres: Insights from a Logistic Regression Analysis\thanks{Code and data are available at: \url{https://github.com/RohanAlexander/starter_folder}.}

### 2024-12-04 paper/paper.tex
paper with title

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={My title},
  pdfauthor={First author; Another author},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{My title\thanks{Code and data are available at:
\url{https://github.com/RohanAlexander/starter_folder}.}}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{My subtitle if needed}
\author{First author \and Another author}
\date{December 3, 2024}

\begin{document}
\maketitle
\begin{abstract}
First sentence. Second sentence. Third sentence. Fourth sentence.
\end{abstract}

@@

To investigate the factors influencing subsidy allocation to licensed
child care centres, a reduced logistic regression model was specified.
The dependent variable,\texttt{subsidy}, is a binary indicator
representing whether a child care centre receives government subsidy (1
= subsidized, 0 = not subsidized). The model includes three key
predictors: \texttt{Operating\ Auspice\ (AUSPICE)}: A categorical
variable indicating the governance model of the child care centre (e.g.,
Non-Profit Agency, Other). Non-profit agencies are hypothesized to be
positively associated with subsidy allocation, consistent with previous
research emphasizing their prioritization in funding schemes (Cleveland
\& Krashinsky, 2005).
\texttt{Participation\ in\ the\ CWELCC\ Program\ (cwelcc\_flag)}: A
binary variable capturing whether the centre participates in the
Canada-Wide Early Learning and Child Care program (1 = participates, 0 =
does not participate). Centres participating in this initiative are
expected to have higher odds of receiving subsidies due to their
alignment with government objectives of affordability and accessibility
(Friendly \& Ballantyne, 2022). \texttt{Total\ Space\ (TOTSPACE)}: A
continuous variable representing the number of licensed spaces available
in the centre. Larger centres are hypothesized to have higher odds of
subsidy allocation, as they can accommodate more families and align with
policy goals of maximizing access (Forry et al., 2010). The logistic
regression model can be expressed mathematically as follows:

@@

The regression model reveals several key observations regarding the
predictors and overall goodness of fit. The intercept, estimated at
-4.967, represents the baseline log-odds of the outcome when all
predictors are at their reference or zero level. While not directly
interpretable, it serves as a baseline reference for the model. Among
the predictors, the AUSPICE: Non-Profit Agency category significantly
increases the log-odds of the outcome, with an estimate of 2.883 and a
small standard error of 0.234, indicating a strong and reliable positive
effect. However, the AUSPICE: Other category shows an unusually large
coefficient (18.772) paired with a very high standard error (686.847),
suggesting instability. The CWELCC Flag variable also exhibits a strong
and reliable positive effect, with an estimate of 3.156 and a standard
error of 0.436, indicating that being flagged as CWELCC significantly
increases the log-odds of the outcome. Additionally, the TOTSPACE
variable has a small but statistically significant effect, with an
estimate of 0.014 and a low standard error of 0.003, reflecting
robustness.

@@

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Extract coefficients and confidence intervals}
\NormalTok{tidy\_model }\OtherTok{\textless{}{-}}\NormalTok{ broom}\SpecialCharTok{::}\FunctionTok{tidy}\NormalTok{(reduced\_model\_2, }\AttributeTok{conf.int =} \ConstantTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
Warning: glm.fit: fitted probabilities numerically 0 or 1 occurred
Warning: glm.fit: fitted probabilities numerically 0 or 1 occurred
Warning: glm.fit: fitted probabilities numerically 0 or 1 occurred
Warning: glm.fit: fitted probabilities numerically 0 or 1 occurred
Warning: glm.fit: fitted probabilities numerically 0 or 1 occurred
Warning: glm.fit: fitted probabilities numerically 0 or 1 occurred
Warning: glm.fit: fitted probabilities numerically 0 or 1 occurred
Warning: glm.fit: fitted probabilities numerically 0 or 1 occurred
Warning: glm.fit: fitted probabilities numerically 0 or 1 occurred
Warning: glm.fit: fitted probabilities numerically 0 or 1 occurred
Warning: glm.fit: fitted probabilities numerically 0 or 1 occurred
Warning: glm.fit: fitted probabilities numerically 0 or 1 occurred
Warning: glm.fit: fitted probabilities numerically 0 or 1 occurred
Warning: glm.fit: fitted probabilities numerically 0 or 1 occurred
Warning: glm.fit: fitted probabilities numerically 0 or 1 occurred
Warning: glm.fit: fitted probabilities numerically 0 or 1 occurred
Warning: glm.fit: fitted probabilities numerically 0 or 1 occurred
Warning: glm.fit: fitted probabilities numerically 0 or 1 occurred
Warning: glm.fit: fitted probabilities numerically 0 or 1 occurred
Warning: glm.fit: fitted probabilities numerically 0 or 1 occurred
Warning: glm.fit: fitted probabilities numerically 0 or 1 occurred
Warning: glm.fit: fitted probabilities numerically 0 or 1 occurred
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Highlight significant coefficients}
\NormalTok{tidy\_model }\OtherTok{\textless{}{-}}\NormalTok{ tidy\_model }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{mutate}\NormalTok{(}\AttributeTok{significant =} \FunctionTok{ifelse}\NormalTok{(conf.low }\SpecialCharTok{\textgreater{}} \DecValTok{0} \SpecialCharTok{|}\NormalTok{ conf.high }\SpecialCharTok{\textless{}} \DecValTok{0}\NormalTok{, }\StringTok{"Significant"}\NormalTok{, }\StringTok{"Not Significant"}\NormalTok{))}

\CommentTok{\# Plot with enhancements}
\FunctionTok{ggplot}\NormalTok{(tidy\_model, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ estimate, }\AttributeTok{y =} \FunctionTok{reorder}\NormalTok{(term, }\FunctionTok{abs}\NormalTok{(estimate)), }\AttributeTok{color =}\NormalTok{ significant)) }\SpecialCharTok{+}
  \FunctionTok{geom\_point}\NormalTok{() }\SpecialCharTok{+}
  \FunctionTok{geom\_errorbarh}\NormalTok{(}\FunctionTok{aes}\NormalTok{(}\AttributeTok{xmin =}\NormalTok{ conf.low, }\AttributeTok{xmax =}\NormalTok{ conf.high), }\AttributeTok{height =} \FloatTok{0.2}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{scale\_color\_manual}\NormalTok{(}\AttributeTok{values =} \FunctionTok{c}\NormalTok{(}\StringTok{"Significant"} \OtherTok{=} \StringTok{"red"}\NormalTok{, }\StringTok{"Not Significant"} \OtherTok{=} \StringTok{"black"}\NormalTok{)) }\SpecialCharTok{+}
  \FunctionTok{labs}\NormalTok{(}
    \AttributeTok{title =} \StringTok{"Coefficient Plot"}\NormalTok{,}
    \AttributeTok{x =} \StringTok{"Coefficient Estimate"}\NormalTok{,}
    \AttributeTok{y =} \StringTok{"Predictor Variables"}\NormalTok{,}
    \AttributeTok{color =} \StringTok{"Significance"}
\NormalTok{  ) }\SpecialCharTok{+}
  \FunctionTok{theme\_minimal}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\includegraphics{paper_files/figure-pdf/unnamed-chunk-16-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Calculate log{-}likelihood of the fitted model}
\NormalTok{log\_likelihood\_fitted }\OtherTok{\textless{}{-}} \FunctionTok{logLik}\NormalTok{(reduced\_model\_2)}
\CommentTok{\# Fit the null model (model with intercept only)}
\NormalTok{null\_model }\OtherTok{\textless{}{-}} \FunctionTok{glm}\NormalTok{(subsidy }\SpecialCharTok{\textasciitilde{}} \DecValTok{1}\NormalTok{, }\AttributeTok{data =}\NormalTok{ train\_data, }\AttributeTok{family =}\NormalTok{ binomial)}
\NormalTok{log\_likelihood\_null }\OtherTok{\textless{}{-}} \FunctionTok{logLik}\NormalTok{(null\_model)}
\CommentTok{\# McFadden\textquotesingle{}s R{-}squared}
\NormalTok{mcfadden\_r2 }\OtherTok{\textless{}{-}} \DecValTok{1} \SpecialCharTok{{-}} \FunctionTok{as.numeric}\NormalTok{(log\_likelihood\_fitted }\SpecialCharTok{/}\NormalTok{ log\_likelihood\_null)}
\CommentTok{\# Print the result}
\FunctionTok{print}\NormalTok{(}\FunctionTok{paste}\NormalTok{(}\StringTok{"McFadden\textquotesingle{}s R{-}squared:"}\NormalTok{, }\FunctionTok{round}\NormalTok{(mcfadden\_r2, }\DecValTok{6}\NormalTok{)))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
[1] "McFadden's R-squared: 0.454164"
\end{verbatim}

@@

The ROC curve in Figure~\ref{fig-roc} shown further validates the
model's classification ability, with an AUC of 0.8998. An AUC value
close to 1 indicates excellent discriminative power, where the model
effectively separates true positives from false positives. The AUC of
0.8998 places this model on the borderline of very good and excellent
discrimination, underscoring its robust predictive capacity. These
results collectively suggest that the logistic regression model not only
fits the data well but also performs strongly in classification tasks.
However, as suggested by Harrell (Harrell 2012), care must be taken to
ensure the model is not overfitted, particularly given the high AUC
value. Cross-validation or testing on an independent dataset is
recommended to confirm the model's generalizability. Additionally,
assessing the contribution of individual predictors can enhance the
interpretability of the model, as overly complex models may inflate
performance metrics without offering meaningful insights. Overall, the
model's performance metrics strongly support its use for prediction in
practical settings while aligning with academic standards for model
evaluation.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Generate predicted probabilities for the test dataset}
\NormalTok{test\_data}\SpecialCharTok{$}\NormalTok{predicted\_prob }\OtherTok{\textless{}{-}} \FunctionTok{predict}\NormalTok{(reduced\_model\_2, }\AttributeTok{newdata =}\NormalTok{ test\_data, }\AttributeTok{type =} \StringTok{"response"}\NormalTok{)}

\CommentTok{\# Predicted vs. Observed Scatterplot}
\FunctionTok{ggplot}\NormalTok{(test\_data, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x =}\NormalTok{ predicted\_prob, }\AttributeTok{y =}\NormalTok{ subsidy)) }\SpecialCharTok{+}
  \FunctionTok{geom\_jitter}\NormalTok{(}\AttributeTok{width =} \FloatTok{0.02}\NormalTok{, }\AttributeTok{height =} \FloatTok{0.02}\NormalTok{, }\AttributeTok{alpha =} \FloatTok{0.6}\NormalTok{, }\AttributeTok{color =} \StringTok{"blue"}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{geom\_smooth}\NormalTok{(}\AttributeTok{method =} \StringTok{"loess"}\NormalTok{, }\AttributeTok{color =} \StringTok{"red"}\NormalTok{, }\AttributeTok{se =} \ConstantTok{FALSE}\NormalTok{) }\SpecialCharTok{+}
  \FunctionTok{labs}\NormalTok{(}
    \AttributeTok{title =} \StringTok{"Predicted vs. Observed Probability"}\NormalTok{,}
    \AttributeTok{x =} \StringTok{"Predicted Probability"}\NormalTok{,}
    \AttributeTok{y =} \StringTok{"Observed Outcome"}
\NormalTok{  ) }\SpecialCharTok{+}
  \FunctionTok{theme\_minimal}\NormalTok{()}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
`geom_smooth()` using formula = 'y ~ x'
\end{verbatim}

\includegraphics{paper_files/figure-pdf/unnamed-chunk-19-1.pdf}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Create a coefficient table}
\NormalTok{model\_list }\OtherTok{\textless{}{-}} \FunctionTok{list}\NormalTok{(}
  \StringTok{"Full Model"} \OtherTok{=}\NormalTok{ log\_model,}
  \StringTok{"AIC Model"} \OtherTok{=}\NormalTok{ reduced\_model\_re,}
  \StringTok{"Simplified Model"} \OtherTok{=}\NormalTok{ reduced\_model\_2}
\NormalTok{)}

\CommentTok{\# Customize the table to display only AIC and Deviance}
\FunctionTok{modelsummary}\NormalTok{(}
\NormalTok{  model\_list,}
  \AttributeTok{output =} \StringTok{"markdown"}\NormalTok{,}
  \AttributeTok{gof\_map =} \FunctionTok{c}\NormalTok{(}\StringTok{"AIC"}\NormalTok{, }\StringTok{"Deviance"}\NormalTok{)  }\CommentTok{\# Include only AIC and Deviance in the table}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}
\centering
\begin{tblr}[         %% tabularray outer open
]                     %% tabularray outer close
{                     %% tabularray inner open
colspec={Q[]Q[]Q[]Q[]},
column{1}={halign=l,},
column{2}={halign=c,},
column{3}={halign=c,},
column{4}={halign=c,},
}                     %% tabularray inner close
\toprule
& Full Model & AIC Model & Simplified Model \\ \midrule %% TinyTableHeader
(Intercept)                               & 10.717     & 10.655     & -4.967    \\
& (2454.170) & (2457.796) & (0.496)   \\
ward                                      & -0.007     &            &           \\
& (0.021)    &            &           \\
AUSPICENon Profit Agency                  & 2.752      & 2.753      & 2.883     \\
& (0.339)    & (0.339)    & (0.234)   \\
AUSPICEOther                              & 19.252     & 19.261     & 18.772    \\
& (1009.264) & (1008.865) & (686.847) \\
bldg_typeCatholic Elementary School       & -14.928    & -14.930    &           \\
& (2454.170) & (2457.796) &           \\
bldg_typeCatholic High School             & 1.120      & 1.037      &           \\
& (6969.058) & (6970.335) &           \\
bldg_typeChurch                           & -16.826    & -16.826    &           \\
& (2454.170) & (2457.796) &           \\
bldg_typeCommercial Building              & -18.590    & -18.616    &           \\
& (2454.171) & (2457.796) &           \\
bldg_typeCommunity College/University     & 1.009      & 0.989      &           \\
& (4488.740) & (4490.254) &           \\
bldg_typeCommunity Health Centre          & 1.092      & 1.037      &           \\
& (6969.058) & (6970.335) &           \\
bldg_typeCommunity Rec/Centre - Board Run & -35.148    & -35.072    &           \\
& (6969.058) & (6970.335) &           \\
bldg_typeCommunity/Rec Centre - City      & 1.053      & 1.077      &           \\
& (4061.162) & (4063.980) &           \\
bldg_typeCommunity/Recreation Centre      & -16.003    & -16.013    &           \\
& (2454.170) & (2457.796) &           \\
bldg_typeHigh Rise Apartment              & -12.625    & -12.631    &           \\
& (2454.170) & (2457.796) &           \\
bldg_typeHospital/Health Centre           & 1.367      & 1.354      &           \\
& (6969.058) & (6970.335) &           \\
bldg_typeHouse                            & -14.939    & -14.920    &           \\
& (2454.170) & (2457.796) &           \\
bldg_typeIndustrial Building              & -32.542    & -32.470    &           \\
& (4618.831) & (4622.295) &           \\
bldg_typeLow Rise Apartment               & -14.884    & -14.852    &           \\
& (2454.173) & (2457.798) &           \\
bldg_typeOffice Building                  & -16.708    & -16.695    &           \\
& (2454.170) & (2457.796) &           \\
bldg_typeOther                            & -16.973    & -16.963    &           \\
& (2454.170) & (2457.796) &           \\
bldg_typePlace of Worship                 & -16.929    & -16.931    &           \\
& (2454.170) & (2457.796) &           \\
bldg_typePrivate Elementary School        & -34.849    & -34.836    &           \\
& (3381.477) & (3387.732) &           \\
bldg_typePublic (school closed)           & -16.221    & -16.188    &           \\
& (2454.170) & (2457.796) &           \\
bldg_typePublic Elementary School         & -13.716    & -13.712    &           \\
& (2454.170) & (2457.796) &           \\
bldg_typePublic Elementary Special        & -0.010     & 0.007      &           \\
& (6969.058) & (6970.335) &           \\
bldg_typePublic High School               & -15.904    & -15.871    &           \\
& (2454.170) & (2457.796) &           \\
bldg_typePublic Middle School             & 1.045      & 1.117      &           \\
& (6969.058) & (6970.335) &           \\
bldg_typePurpose Built                    & -14.567    & -14.572    &           \\
& (2454.170) & (2457.796) &           \\
bldg_typeSynagogue                        & -18.451    & -18.405    &           \\
& (2454.170) & (2457.796) &           \\
cwelcc_flag                               & 3.504      & 3.473      & 3.156     \\
& (0.519)    & (0.510)    & (0.436)   \\
TOTSPACE                                  & 0.013      & 0.013      & 0.014     \\
& (0.004)    & (0.004)    & (0.003)   \\
\bottomrule
\end{tblr}
\end{table}

>>>>

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Understanding Subsidy Allocation in Toronto's Child Care Centres: Insights from a Logistic Regression Analysis},
  pdfauthor={Claire Ma},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Understanding Subsidy Allocation in Toronto's Child Care Centres:
Insights from a Logistic Regression Analysis\thanks{Code and data are
available at: \url{https://github.com/RohanAlexander/starter_folder}.}}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{Non-Profit Governance, Program Participation, and Capacity
Drive Funding Decisions}
\author{Claire Ma}
\date{December 3, 2024}

\begin{document}
\maketitle
\begin{abstract}
This study examines the factors influencing government subsidy
allocation to licensed child care centres in Toronto based on Licened
child Care Centers from Open Data Toronto. By employing a logistic
regression model, we found that non-profit governance, participation in
the Canada-Wide Early Learning and Child Care (CWELCC) program, and
larger capacity significantly increase a centre's likelihood of
receiving subsidies. These findings highlight how funding priorities
align with public goals to expand access to high-quality child care,
especially for vulnerable populations. By uncovering patterns in subsidy
distribution, this research provides critical insights for policymakers
to ensure more equitable and effective resource allocation.
\end{abstract}

@@

To investigate the factors influencing subsidy allocation to licensed
child care centres, a reduced logistic regression model was specified.
The dependent variable,\texttt{subsidy}, is a binary indicator
representing whether a child care centre receives government subsidy (1
= subsidized, 0 = not subsidized). The model includes three key
predictors: \texttt{Operating\ Auspice\ (AUSPICE)}: A categorical
variable indicating the governance model of the child care centre (e.g.,
Non-Profit Agency, Other). Non-profit agencies are hypothesized to be
positively associated with subsidy allocation, consistent with previous
research emphasizing their prioritization in funding schemes (Cleveland
\& Krashinsky, 2005).
\texttt{Participation\ in\ the\ CWELCC\ Program\ (cwelcc\_flag)}: A
binary variable capturing whether the centre participates in the
Canada-Wide Early Learning and Child Care program (1 = participates, 0 =
does not participate). Centres participating in this initiative are
expected to have higher odds of receiving subsidies due to their
alignment with government objectives of affordability and accessibility
(Friendly \& Ballantyne, 2022). \texttt{Total\ Space\ (TOTSPACE)}: A
continuous variable representing the number of licensed spaces available
in the centre. Larger centres are hypothesized to have higher odds of
subsidy allocation, as they can accommodate more families and align with
policy goals of maximizing access (Forry et al., 2010). The logistic
regression model can be expressed mathematically as follows: {[}
\text{logit}(P(Y=1)) = \beta\_0 + \beta\_1 (\text{Non-Profit Auspice}) +
\beta\_2 (\text{CWELCC Participation}) + \beta\_3 (\text{Total Space}) +
\epsilon {]}

Where:

\begin{itemize}
    \item \( P(Y=1) \): The probability that a child care center receives a subsidy (\( Y=1 \)).
    \item \( \text{logit}(P(Y=1)) \): The log-odds of receiving a subsidy, defined as:
    \[
    \text{logit}(P(Y=1)) = \ln\left(\frac{P(Y=1)}{1 - P(Y=1)}\right)
    \]
    This transformation ensures that the predicted probabilities remain between 0 and 1.
\end{itemize}

\subsubsection{Parameters:}\label{parameters}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  **( \beta\_0 ) (Intercept):**

  \begin{itemize}
  \tightlist
  \item
    Represents the baseline log-odds of subsidy allocation when all
    predictors are set to their reference categories or zero values.
  \item
    It provides the starting point for predicting subsidy probabilities.
  \end{itemize}
\item
  **( \beta\_1 (\text{Non-Profit Auspice}) ):**

  \begin{itemize}
  \tightlist
  \item
    Measures the effect of a center being operated as a non-profit
    compared to the baseline (e.g., commercial auspice) on the log-odds
    of subsidy allocation.
  \item
    A positive ( \beta\_1 ) increases the likelihood of receiving a
    subsidy for non-profit centers relative to the reference group.
  \end{itemize}
\item
  **( \beta\_2 (\text{CWELCC Participation}) ):**

  \begin{itemize}
  \tightlist
  \item
    Captures the influence of participating in the Canada-Wide Early
    Learning and Child Care (CWELCC) program on the log-odds of subsidy
    allocation.
  \end{itemize}
\item
  **( \beta\_3 (\text{Total Space}) ):**

  \begin{itemize}
  \tightlist
  \item
    Represents the effect of the number of licensed child care spaces (a
    continuous variable) on the log-odds of subsidy allocation.
  \end{itemize}
\item
  \textbf{( \epsilon ) (Error Term):}

  \begin{itemize}
  \tightlist
  \item
    Accounts for unexplained variation in the log-odds of subsidy
    allocation not captured by the predictors.
  \end{itemize}
\end{enumerate}

@@

The regression model reveals several key observations regarding the
predictors and overall goodness of fit. The intercept, estimated at
-4.967, represents the baseline log-odds of the outcome when all
predictors are at their reference or zero level. While not directly
interpretable, it serves as a baseline reference for the model. Among
the predictors, the AUSPICE: Non-Profit Agency category significantly
increases the log-odds of the outcome, with an estimate of 2.883 and a
small standard error of 0.234, indicating a strong and reliable positive
effect. However, the AUSPICE: Other category shows an unusually large
coefficient (18.772) paired with a very high standard error (686.847),
suggesting instability. The CWELCC Flag variable also exhibits a strong
and reliable positive effect, with an estimate of 3.156 and a standard
error of 0.436, indicating that being flagged as CWELCC significantly
increases the log-odds of the outcome. Additionally, the TOTSPACE
variable has a small but statistically significant effect, with an
estimate of 0.014 and a low standard error of 0.003, reflecting
robustness. Figure~\ref{fig-cp} further visualizes the significance of
predictors.

\includegraphics{paper_files/figure-pdf/fig-cp-1.pdf}

\caption{\label{fig-cp}Coefficient Plot}

@@

@@

The ROC curve in Figure~\ref{fig-roc} shown further validates the
model's classification ability, with an AUC of 0.8998. An AUC value
close to 1 indicates excellent discriminative power, where the model
effectively separates true positives from false positives. The AUC of
0.8998 places this model on the borderline of very good and excellent
discrimination, underscoring its robust predictive capacity.

\includegraphics{paper_files/figure-pdf/fig-pvo-1.pdf}

\caption{\label{fig-pvo}Predicted vs.~Observed Probability}

Figure~\ref{fig-pvo} provides a visual representation of the alignment
between the predicted probabilities generated by the logistic regression
model and the actual observed binary outcomes. This diagnostic tool is
crucial for assessing the model's calibration and predictive accuracy,
as emphasized by Harrell (Harrell 2012). The x-axis shows the predicted
probabilities ranging from 0 to 1, while the y-axis represents the
observed outcomes, where 0 indicates the absence of an event and 1
indicates its presence. The red diagonal line serves as the reference
for perfect calibration, where predicted probabilities perfectly match
observed outcomes. The clustering of blue points near 0 and 1 along the
y-axis indicates that the model effectively distinguishes between the
two classes, a key goal in binary classification modeling. This
observation aligns with the high AUC value of 0.8998 observed in the ROC
curve, which demonstrates the model's excellent discriminative ability.
Additionally, points near the diagonal line further suggest strong
calibration, where the predicted probabilities closely align with actual
outcomes. The concentration of blue points at the extremes of 0 and 1
highlights that the model makes confident and accurate predictions in
many cases However, some dispersion is observed at lower predicted
probabilities, particularly around 0.25 and 0.5. These deviations may
reflect cases where the model struggles to classify outcomes or makes
less confident predictions. Such discrepancies could indicate areas
where further refinement of the model or predictors is necessary. These
deviations could also be linked to specific subgroups or predictor
interactions, necessitating additional diagnostics or recalibration, as
suggested by Steyerberg et al. (Steyerberg et al. 2010). Overall, this
graph demonstrates that the model is well-calibrated and effective in
its predictions, with minor limitations at lower probability levels. The
close alignment of predictions with observed outcomes supports the
model's reliability in classification tasks. This interpretation aligns
with other metrics, such as McFadden's R2R2
(\textbf{mcfadden1972conditional?}) and AUC, which indicate robust
overall model performance while pointing to opportunities for
refinement. Including this visualization in the results section provides
a clear and data-driven assessment of the model's strengths and areas
for potential improvement.

\#Discussion

Subsidy allocation to licensed child care centres in Toronto addresses
the critical issue of affordability, making high-quality early childhood
education accessible to more families as licensed centres must meet
strict safety, staffing, and programming standards, which foster
children's cognitive, social, and emotional growth (Cleveland and
Krashinsky 2009). Without subsidies, the cost of regulated care can be
prohibitive, particularly for low-income households. This paper
undertakes a detailed examination of the factors influencing subsidy
allocation to licensed child care centres in Toronto, leveraging the
City of Toronto's Licensed Child Care Centres dataset from Open Data
porta, which includes detailed information about 1,071 licensed
centers(Gelfand 2022). This study employs a logistic regression
model-binomial family to identify relationships between key predictor
variables and the binary dependent variable -- receiving subsidies.

The dataset underwent meticulous preprocessing to ensure reliability and
accuracy for analysis. This involved several key steps of data cleaning,
variable transformation and training-test Split. Initially, variables of
Operating auspice(AUSPICE), ward, building type(bldg\_type), total
space(TOTSPACE) and participation status in programs(cwelcc\_flag) were
chosen based their relevance and practical influence subsidy outcomes
and forecasting supported by scholar literatures. After fitting and
improving the model using stepwise selection based on the Akaike
Information Criterion (AIC), non-significant variables (ward and
building type(bldg\_type)) were removed. The final model retained
Operating Auspice, CWELCC Participation, and Total Licensed Capacity as
key predictors. The assumptions underpinning the logistic regression
model were thoroughly validated, ensuring its statistical rigor and
reliability. Independence of observations was confirmed, as each data
point corresponds to a unique child care centre, eliminating concerns of
clustering or dependency. Linearity between continuous predictors, such
as Total Licensed Capacity, and the logit transformation was validated
through component + residual plots, which revealed consistent patterns,
supporting the appropriateness of the model. Additionally,
multicollinearity was minimized, with Variance Inflation Factor (VIF)
values well below the critical threshold of 5, indicating that the
predictors are independent and the coefficient estimates are stable.
Together, these validations affirm the robustness of the model and its
suitability for the analysis.

The result coefficients in table the logistic regression model provide a
detailed understanding of how key predictors influence the likelihood of
subsidy allocation to licensed child care centres in Toronto. By
measuring the change in log-odds of subsidy allocation associated with
each predictor, the model quantifies the relative importance of
governance models, program participation, and operational capacity. The
coefficient for Non-Profit Auspice emerged as strongly positive and
statistically significant, indicating that non-profit child care centres
are far more likely to receive subsidies than commercial or public
centres. This is reflective of policy priorities that favor
organizations with a social mission over those with profit-oriented
objectives. Non-profits are often viewed as better aligned with public
goals of equity and accessibility, as they reinvest surplus revenues
into improving care quality rather than distributing profits
(\textbf{hansmann1980role?}). This is consistent with research by
(\textbf{salamon1995partners?}), who noted that non-profits often fill
critical gaps in the provision of public services, making them natural
candidates for targeted funding. In the context of child care,
non-profits may also be more likely to serve economically vulnerable
populations, further justifying their prioritization for subsidies. The
odds ratio derived from this coefficient underscores the strength of
this relationship, suggesting that non-profit centres are several times
more likely to receive subsidies compared to commercial centres.
Participation in the Canada-Wide Early Learning and Child Care (CWELCC)
program was the most significant predictor in the model, with a highly
positive coefficient. This result highlights the critical role of policy
alignment in subsidy allocation. Centres participating in CWELCC
demonstrate compliance with national and provincial affordability
initiatives, making them high-priority recipients of government funding.
Research by (\textbf{bennett2011working?}) supports this finding,
emphasizing that early childhood programs aligned with broader
governmental objectives are often better positioned to secure resources.
The odds ratio for CWELCC participation indicates that such centres are
substantially more likely to receive subsidies, which is consistent with
targeted funding strategies aimed at expanding access to high-quality,
affordable child care (\textbf{kaga2010caring?}). Total Licensed
Capacity (TOTSPACE), representing the number of spaces a centre is
licensed to operate, was another important predictor, although its
coefficient was smaller in magnitude compared to governance and CWELCC
participation. This positive coefficient suggests that larger centres
are more likely to receive subsidies, likely due to their capacity to
serve more families and their operational scale. Larger centres can
achieve economies of scale and often have greater visibility, making
them attractive candidates for public funding
(\textbf{weisbrod1998profit?}). However, the smaller effect size of this
variable indicates that capacity alone is insufficient to determine
subsidy allocation. This finding aligns with
(\textbf{penn2011quality?}), who argues that while capacity is an
important practical consideration, other factors---such as governance
and programmatic alignment---carry more weight in funding decisions.
Interpreting these coefficients through odds ratios further clarifies
their impact. For instance, the odds ratio for CWELCC participation
suggests that centres in the program are several times more likely to
receive subsidies compared to non-participating centres, holding other
factors constant. Similarly, the odds ratio for non-profit governance
reinforces the preferential treatment of non-profits in subsidy
allocation policies. These findings align with the broader literature on
public funding, which emphasizes the strategic use of subsidies to
support centres that meet both operational and policy criteria
(\textbf{penn2011quality?}). The statistical significance of these
coefficients underscores their reliability as shown in figure. Variables
such as Non-Profit Auspice and CWELCC Participation demonstrated small
standard errors and strong significance levels, reinforcing their
influence on subsidy allocation. In contrast, variables with less
consistent effects, such as certain building types or governance
categories in the ``Other'' category, exhibited instability, suggesting
that their role in funding decisions may be more context-specific or
secondary. The model's findings highlight a strategic approach to
subsidy allocation, where funding decisions are guided by alignment with
policy goals, operational capacity, and governance structures.

The logistic regression model applied in this research proves to be an
ideal tool for analyzing subsidy allocation to licensed child care
centres in Toronto, particularly due to its capacity to handle binary
outcomes effectively. By modeling the log-odds of an event, logistic
regression ensures predicted probabilities remain between 0 and 1,
aligning seamlessly with the binary nature of the dependent
variable---whether a centre receives a subsidy. This feature makes the
model both statistically appropriate and easy to interpret. A notable
strength of the model lies in its interpretability. Coefficients can be
transformed into odds ratios, providing a clear understanding of how
predictors influence the likelihood of receiving subsidies. For example,
the positive coefficients for Non-Profit Auspice and CWELCC
Participation underscore their significant, policy-relevant impact on
subsidy distribution. This clarity makes the model particularly valuable
for stakeholders and policymakers who require actionable and easily
communicated insights. The model's performance metrics further
demonstrate its effectiveness. A McFadden's R² value of 0.454 indicates
that the model explains a substantial portion of the variance in subsidy
allocation. Additionally, the Area Under the Curve (AUC) of 0.8998
highlights the model's exceptional discriminative ability, confirming
its capacity to accurately differentiate between subsidized and
non-subsidized centres. The low Root Mean Squared Error (RMSE) of 0.32
further attests to the model's strong predictive performance. Together,
these metrics validate the model's reliability and robustness. Another
key advantage of logistic regression is its simplicity compared to more
complex approaches such as random forests or decision trees. While those
alternatives might offer marginally higher predictive power, they lack
the transparency and interpretability required for policy-focused
research. Logistic regression strikes a critical balance between
accuracy and clarity, making it a practical choice for guiding equitable
subsidy allocation strategies. Moreover, the flexibility of logistic
regression allows for extensions to explore more intricate
relationships. For instance, incorporating interaction terms or temporal
data could provide deeper insights while maintaining the model's
usability. This adaptability ensures the model remains relevant and
applicable as policy environments evolve.

Our model, while effective, has several limitations that warrant
consideration. One of the primary limitations is its reliance on the
assumption of linearity between predictors and the log-odds of the
outcome. Although this assumption is often reasonable, it may
oversimplify complex, non-linear relationships that can exist in
real-world phenomena such as subsidy allocation. As
(\textbf{harrell2015regression?}) notes, failing to capture these
non-linear relationships can lead to biased estimates and limit the
model's predictive power. Another limitation is the model's dependence
on correct specification of predictors. Excluding relevant variables or
including irrelevant ones in initial variable selection can result in
omitted variable bias or overfitting, respectively. Additionally, the
model does not inherently account for higher-order interactions unless
explicitly included. For instance, the combined effect of Non-Profit
Auspice and CWELCC Participation could provide valuable insights into
how policy and governance jointly influence subsidy allocation. (Hosmer
Jr, Lemeshow, and Sturdivant 2013) highlight that logistic regression
can oversimplify complex relationships if interaction terms are not
considered, potentially leading to incomplete conclusions. Finally, the
static nature of the model is another drawback. Logistic regression
provides a snapshot of relationships at a single point in time, failing
to capture temporal dynamics or policy changes that can influence
subsidy allocation over time. As Singer and
(\textbf{singer2003applied?}) suggest, incorporating longitudinal data
could provide a more nuanced understanding of how these relationships
evolve.
